---
title: "Big Data project: Car evaluation"
authors: "Alexandra DU - Thibaud VINANT - Chenyu YING - Loïc NEVOUET


output:
  html_document: default
  pdf_document: default
  word_document: default
---


```{r}
library(ggplot2)
library(readxl)
library(naivebayes)
library(plotly)
library(shiny)
library(reshape)
library(lda)
library(MASS)
library(class)
library(gplots)
library(gmodels)
ggplot<-ggplot2::ggplot
```

Source of data set:

http://archive.ics.uci.edu/ml/datasets/Car+Evaluation


#<font color="blue">1) Data preparation</font>

<h2><font color="teal">1.1) Data importation and preliminary visualization</font></h2>

```{r}
data<- read.table("car.data", sep=",")
colnames(data) <- c("Buying_price", "Maintenance_price", "Doors", "Persons", "Lug_boot", "Safety","Value")

#Reorder the variable scales from low to high
data$Buying_price<-factor(data$Buying_price, levels=c("low","med","high","vhigh"))
data$Maintenance_price<-factor(data$Maintenance_price, levels=c("low","med","high","vhigh"))
data$Value<-factor(data$Value, levels=c("unacc","acc","good","vgood"))
data$Lug_boot<-factor(data$Lug_boot, levels=c("small","med","big"))
data$Safety<-factor(data$Safety, levels=c("low","med","high"))
```

Description of the variables: 

- Buying price and maintenance price: low, med, high, very high

- Number of doors: 2, 3, 4, 5 or more

- Max number of people: 2, 4, or more

- Size of luggage boot: small, medium, big

- Safety index: low, medium, high

- Value of car: what we are trying to predict. Unacceptable, acceptable, good, very good

```{r}
#Roughly explore the influence of each parameter on the Value
graph_safety=ggplot(data,aes(x=Safety,y=Value))+geom_point(position = "jitter",aes(color=Value))
graph_safety

graph_buying=ggplot(data,aes(x=Buying_price,y=Value))+geom_point(position = "jitter",aes(color=Value))
graph_buying

graph_maint=ggplot(data,aes(x=Maintenance_price,y=Value))+geom_point(position = "jitter",aes(color=Value))
graph_maint

graph_lug=ggplot(data,aes(x=Lug_boot,y=Value))+geom_point(position = "jitter",aes(color=Value))
graph_lug

graph_persons=ggplot(data,aes(x=Persons,y=Value))+geom_point(position = "jitter",aes(color=Value))
graph_persons

graph_doors=ggplot(data,aes(x=Doors,y=Value))+geom_point(position = "jitter",aes(color=Value))
graph_doors
```
Temporary conclusions: 

- Safety is a quite important factor: a low safety automatically results in "Unacceptable"
- Buying price: if high or very high, results in "unacc" or "acc" at best
- Maintenance price: if very high, results in "unacc" or "acc", but some cars with "high" maintenance price (sign of a good quality?) are "very good"
- Number of doors: irrelevant in determining value at first glance because the proportions of unacc, acc, good and vgood seem to not depend on the number of doors
- Max number of passengers: "2 people" means a car that is too small --> unacc. Indifferent between 4 people or more
- Size of luggage loot: not very relevant but a small loot cannot result in "very good"


<h2><font color="teal">1.2) Data cleaning</font></h2>

We removed the data that seemed illogical, namely the cars that have 4 or 5 doors but only a capacity of 2 people.
As for the separation of the levels of some variables into new variables (such as Doors2 that contains a boolean according to whether or not a car has 2 doors), it is justified by our observations when running the logistic regression (see part 2).
```{r}
#data preparation

data2 <-subset(data, (data$Doors!=4 | data$Persons!=2) & (data$Doors!="5more" | data$Persons!=2)) #data cleaning for inconsistent results

#reorganization of data
data2$Doors2<-(data2$Doors==2)
data2$Doors3<-(data2$Doors==3)
data2$Doors4<-(data2$Doors==4)
data2$Doors5<-(data2$Doors=="5more")

data2$Buying_price_low<-(data2$Buying_price=="low")
data2$Buying_price_med<-(data2$Buying_price=="med")
data2$Buying_price_high<-(data2$Buying_price=="high")
data2$Buying_price_vhigh<-(data2$Buying_price=="vhigh")

data2$Maintenance_price_low<-(data2$Maintenance_price=="low")
data2$Maintenance_price_med<-(data2$Maintenance_price=="med")
data2$Maintenance_price_high<-(data2$Maintenance_price=="high")
data2$Maintenance_price_vhigh<-(data2$Maintenance_price=="vhigh")

data2$value2<-as.numeric(data2$Value)
```

<h2><font color="teal">1.3) Visualization of relations between attributes</font></h2>

```{r}
#Correlation matrix with the initial categories
data3 <-subset(data, (data$Doors!=4 | data$Persons!=2) & (data$Doors!="5more" | data$Persons!=2))

numerica<-data.frame(as.numeric(data3$Buying_price),as.numeric(data3$Maintenance_price),as.numeric(data3$Doors),as.numeric(data3$Persons),as.numeric(data3$Safety),as.numeric(data3$Lug_boot))
cormat.test<-round(cor(numerica),2)
head(cormat.test)

melted_cormat.test <- melt(cormat.test)
head(melted_cormat.test)

ggplot(data = melted_cormat.test, aes(x=X1, y=X2, fill=value)) + 
  geom_tile()
```

We see that - as we could expect - after the data cleaning, there is a correlation between the number of doors and the number of persons that can fit in.

#<font color="blue">2) Method 1: Logistic regression</font>

<h2><font color="teal">2.1) Generate randomly the training data and the test data</font></h2>

```{r}
set.seed(3)
size.dtrain<-floor(0.55*nrow(data2)) #we take 55% of the data as training data
size.dtest<-nrow(data2)-size.dtrain
index.train <- sample(1:nrow(data2),size.dtrain,replace=FALSE) 

dtrain<-data2[index.train,]
dtest<-data2[-index.train,] #the remaining 45% are test data

attach(dtrain)
```

<h2><font color="teal">2.2) Model that includes all parameters in predicting the value</font></h2>





<h3>Conclusions:</h3> 
- The number of people is totally irrelevant (p-value is too large).
- It appears that safety is irrelevant (p-value too large).
- For the same attribute (for instance buying price), some levels (low and medium) have too large p-values whereas others (high and very high) have small enough p-values. In order to eliminate the irrelevant levels, we separated the corresponding levels into new boolean variables (see part 1.2).

<h2><font color="teal">2.3) Model that includes only relevant parameters in predicting the value</font></h2>

```{r}
glmfit<-glm(Value~ Buying_price_high + Buying_price_vhigh + Maintenance_price_high + Maintenance_price_vhigh +Doors5+ Doors4 + Lug_boot, family=binomial,data=dtrain)

              
proba <- predict(glmfit,type="response",data=dtrain)
summary(proba)

dataframe<-data.frame(proba)

p <- ggplot(dataframe, aes(proba)) + geom_histogram(aes(y=..density..), colour="black", fill="white") + geom_density(alpha=.2, fill="#FF6666")
p

summary(glmfit)

```

We have 4 categories, so we need to set 3 thresholds for the predicted probability. The higher the latter is, the better the car's value is.

```{r}
#Function that returns the vector with associates to each car its predicted Value, depending on the chosen thresholds for the probabilities
result.table<-function(p1,p2,p3,dframe){
  glm.probs=predict(glmfit,type="response",dframe)
  glm.pred = rep(1,nrow(dframe))# creates a vector of 1 elements 
  glm.pred[glm.probs>=p1]=4 #transforms to 1 all of the elements for which th e predicted
  glm.pred[(glm.probs>=p2) & (glm.probs<p1)]=3
  glm.pred[glm.probs>=p3 & glm.probs<p2]=2
  
  return(glm.pred)
}

#Function that optimizes the choice of thresholds to achieve the highest precision:
optimize<-function(p1,p2,p3,interval,dframe){
  results<-list()
  p1bis=p1
  while(p1bis<1){
    p2bis=p2
    while (p2bis<p1bis){
      p3bis=p3
      while (p3bis<p2bis) {
        confusion.matrix=table(result.table(p1bis,p2bis,p3bis,dframe),dframe$value2)
        precision=sum(diag(confusion.matrix))/nrow(dframe)
        results=c(results,list(c(p1bis,p2bis,p3bis,precision)))
        p3bis=p3bis+interval
      }
      p2bis=p2bis+interval
    }
    p1bis=p1bis+interval
  }
  
  d<-as.data.frame(results)
  return(d)
}

q<-optimize(0.3,0.2,0.1,0.05,dtrain)
q<-t(q) #transpose q
q<-as.data.frame(q) #converts to data frame

#Renaming rows and columns
colnames(q)<-c("p1","p2","p3","Precision")
rownames(q)<-c()

#Extracting the maximum precision and its associated threshold probability
qmax<-subset(q, (q$Precision==max(q$Precision)))
qmax
```

```{r}
#Visualize the choice of the optimal thresholds

p <- plot_ly(dtrain, x = ~q$p1, y = ~q$p2, z = ~q$Precision,
        marker = list(color = ~q$p3, colorscale = c('#FFE1A1', '#683531'), showscale = TRUE))
layout(p,title = 'Logistic regression precision as a function of the probability thresholds',
        scene = list(xaxis = list(title = 'p1'),
                     yaxis = list(title = 'p2'),
                     zaxis = list(title = "Model precision")),
       annotations = list(
           x = 1.13,
           y = 1.03,
           text = 'Value of p3',
           xref = 'paper',
           yref = 'paper',
           showarrow = FALSE
         ))
         
```
The red color scale applies to the values of threshold p3.

A maximum precision of 65.7% can be obtained on the training data set, which is better than the naive guess "all cars are unacceptable" (giving a precision of 64.7%).

The confusion matrix associated with the chosen thresholds is:
```{r}
confusion.matrix<-table(result.table(0.8,0.75,0.4,dtrain),dtrain$value2)
confusion.matrix

precision=sum(diag(confusion.matrix))/nrow(dtrain)
precision

heatmap.train=heatmap.2(confusion.matrix,col=rev(heat.colors(100)),xlab="Training_set",ylab="Prediction",cellnote=confusion.matrix,notecol="black",trace="none",dendrogram="none",rownames(confusion.matrix),colnames(confusion.matrix))
```

We compare our result with the proportion of unacc cars in the training set, which is our naive predictor. (Naive guess: "predict all cars are unacceptable") Since this proportion is quite large, the naive guess would actually give a fairly good precision.
```{r}
nrow(subset(dtrain,Value=="unacc"))/nrow(dtrain)
```
The precision of our model is slightly better than the naive guess.

<h2><font color="teal">2.3) Performance of the logistic regression model on the test data set</font></h2>

```{r}
#test data

confusion.matrix.test<-table(result.table(0.8,0.75,0.4,dtest),dtest$value2)
confusion.matrix.test

precision.test=sum(diag(confusion.matrix.test))/nrow(dtest)
precision.test

heatmap.test=heatmap.2(confusion.matrix.test,col=rev(heat.colors(100)),xlab="Test_set",ylab="Prediction",cellnote=confusion.matrix.test,notecol="black",trace="none",dendrogram="none",rownames(confusion.matrix.test),colnames(confusion.matrix.test))

# Compare with the naive guess ie Proportion of unacc cars in test set
nrow(subset(dtest,Value=="unacc"))/nrow(dtest)
```

The model gives a precision of 60.34% on the test set, which is less good than the naive guess, so this method is not optimum.
Particularly, our model does not differenciate well between acceptable and unacceptable cars, and is not very good at finding good/very good cars.

#<font color="blue">3) Method 2: Discriminant analysis</font>

The size of the training set is 792 (55% of the whole data set) and the test set represents 45% of the whole data set.

<h2><font color="teal">3.1) Performance of LDA on the training data set</font></h2>

```{r}
# LDA on training/test set
lda.model1=lda(value2~ Buying_price + Maintenance_price + Safety + Doors + Persons + Lug_boot,data=dtrain)
lda.model1
plot(lda.model1, pch = 20, col=as.integer(dtrain$value))

#Precision on training set
lda.pred0=predict (lda.model1,dtrain)
confusion_matrix_lda_training=table(lda.pred0$class,dtrain$value)
sum(diag(confusion_matrix_lda_training))/nrow(dtrain)
#LDA decomposition on training set
LD1.train=ldahist(lda.pred0$x[,1],g=dtrain$value)
LD2.train=ldahist(lda.pred0$x[,2],g=dtrain$value)
LD3.train=ldahist(lda.pred0$x[,3],g=dtrain$value)
```

On the training data set, the model can reach a precision of 89.9%.

The first linear discriminant (LD1) performs best in separating different groups.


<h2><font color="teal">3.2) Performance of LDA on the test data set</font></h2>

```{r}
#On test set
lda.pred1=predict (lda.model1,dtest)
confusion_matrix_lda_test=table(lda.pred1$class,dtest$value)
confusion_matrix_lda_test
heatmap.lda.test=heatmap.2(confusion_matrix_lda_test,col=rev(heat.colors(100)),xlab="Test set",ylab="Prediction",cellnote=confusion_matrix_lda_test,notecol="black",trace="none",dendrogram="none", rownames(confusion_matrix_lda_test),colnames(confusion_matrix_lda_test))

#Precision on test set
sum(diag(confusion_matrix_lda_test))/nrow(dtest)

```

On the test set, the precision is 86.7%. 

The model enables us to find correctly 90.2% of the unacceptable cars among all unacceptable ones in the test set, and according to the prediction, 98.1% of cars predicted as "unacceptable" are actually "unacceptable".

Relatively, the model is less effective in identifying "acceptable" cars.


#<font color="blue">4) Method 3: K-nearest neighbors</font>

<h2><font color="teal">4.1) Convert data frame into matrix</font></h2>
```{r}
#Conversion of data frames into matrices

Safety.knn=factor(data2$Safety, levels=c("low","med","high"),labels=c("0","1","2"))
Buying_price.knn=factor(data2$Buying_price, levels=c("low","med","high","vhigh"),labels=c("0","1","2","4"))
Maintenance_price.knn=factor(data2$Maintenance_price, levels=c("low","med","high","vhigh"),labels=c("0","1","2","4"))
Doors.knn=factor(data2$Doors, levels=c("2","3","4","5more"),labels=c("0","1","2","3"))
Persons.knn=factor(data2$Persons, levels=c("2","4","more"),labels=c("0","1","2"))
Lug_boot.knn=factor(data2$Lug_boot, levels=c("small","med","big"),labels=c("0","1","2"))
Value.knn=factor(data2$Value, levels=c("unacc","acc","good","vgood"),labels=c("1","2","3","4"))
data.knn=data.frame(Safety.knn,Buying_price.knn,Maintenance_price.knn,Doors.knn,Persons.knn,Lug_boot.knn,Value.knn)
```
The training set and the test set represent respectively 55% and 45% of the whole data set.

```{r}
#Constitute training/test set from the previous matrices

dtrain.knn<-data.knn[index.train,]
dtest.knn<-data.knn[-index.train,]

dtrValue=dtrain.knn[,7]
dteValue=dtest.knn[,7]
```

<h2><font color="teal">4.2) Performance of KNN on the test set </font></h2>

The size of the data set is large enough to involve all 6 attributes as predictors.

```{r}
# 6 attributes

dtrain.knn0=data.frame(dtrain.knn$Buying_price,dtrain.knn$Maintenance_price.knn,dtrain.knn$Doors.knn,dtrain.knn$Lug_boot.knn,dtrain.knn$Persons.knn,dtrain.knn$Safety.knn)
dtest.knn0=data.frame(dtest.knn$Buying_price,dtest.knn$Maintenance_price.knn,dtest.knn$Doors.knn,dtest.knn$Lug_boot.knn,dtest.knn$Persons.knn,dtest.knn$Safety.knn)


dtrValue=dtrain.knn[,7]
dteValue=dtest.knn[,7]

#To find the optimal k, we run a loop
results<-list()
for (k in 1:20) {
  set.seed(1)
  knn.pred0=knn(dtrain.knn0,dtest.knn0,dtrValue,k)
  table.test.knn0=table (knn.pred0,dteValue)
  p=sum(diag(table.test.knn0))/sum(table.test.knn0)
  results=c(results,list(c(k,p)))
  
}
q<-as.data.frame(results)
q<-t(q) #transpose q
q<-as.data.frame(q) #converts to data frame

#Renaming rows and columns
colnames(q)<-c("k","Precision")
rownames(q)<-c()

#Extracting the maximum precision and its associated K
qmax<-subset(q, (q$Precision==max(q$Precision)))
qmax
```
Opimization function gives optimal k=3. We will use this value from now on.

```{r}
set.seed(1)
knn.pred0=knn(dtrain.knn0,dtest.knn0,dtrValue,k=3)
table.test.knn0=table (knn.pred0,dteValue)
table.test.knn0
p0=sum(diag(table.test.knn0))/sum(table.test.knn0)
p0

heatmap=heatmap.2(table.test.knn0,col=rev(heat.colors(100)),xlab="Test_set",ylab="Prediction",cellnote=table.test.knn0,notecol="black",trace="none",dendrogram="none",rownames(table.test.knn0),colnames(table.test.knn0))
CrossTable(dteValue,knn.pred0,prop.chisq = FALSE)
```

When selecting 6 attributes and a k equal to 3, a precision of 92.6% can be reached.

The model is able to identify 94.9% of the unacceptable cars in the test set, which means there is only 5.1% of the possibility that a decision maker would wrongly choose to produce an unsellable car. Regarding the outcome of the prediction, 96.8% of cars are correctly classified as "unacceptable", which means there is a possibility of 3.2% that a decision maker will ignore an "acceptable" car.

The limit of using KNN is that this approach cannot tell how will different single predictor influence the outcome.


<h2><font color="teal">4.3) Some other attempts </font></h2>

We tried some other combinations and other numbers of predictors, but they give less good precisions:

```{r}
# 5 attributes (buying price, maintenance price, doors, luggage boot, safety)
dtrain.knn1=data.frame(dtrain.knn$Buying_price,dtrain.knn$Maintenance_price.knn,dtrain.knn$Doors.knn,dtrain.knn$Lug_boot.knn,dtrain.knn$Safety.knn,dtrain.knn$Value.knn)
dtest.knn1=data.frame(dtest.knn$Buying_price,dtest.knn$Maintenance_price.knn,dtest.knn$Doors.knn,dtest.knn$Lug_boot.knn,dtest.knn$Safety.knn, dtest.knn$Value.knn)
dtrValue1=dtrain.knn1[,6]
dteValue1=dtest.knn1[,6]

results<-list()
for (k in 1:20) {
  set.seed(1)
knn.pred1=knn(dtrain.knn1[,1:5],dtest.knn1[,1:5],dtrValue1,k)
table.test.knn1=table (knn.pred1,dteValue1)
p1=sum(diag(table.test.knn1))/sum(table.test.knn1)
results=c(results,list(c(k,p1)))
  
}
q<-as.data.frame(results)
q<-t(q) #transpose q
q<-as.data.frame(q) #converts to data frame

#Renaming rows and columns
colnames(q)<-c("k","Precision")
rownames(q)<-c()

#Extracting the maximum precision and its associated K
qmax<-subset(q, (q$Precision==max(q$Precision)))
qmax
```
In this case, the highest precision is 84.1%. 

```{r}
# 4 attributes (buying price, Doors, luggage boot,Safety)
dtrain.knn2=data.frame(dtrain.knn$Buying_price,dtrain.knn$Doors.knn,dtrain.knn$Lug_boot.knn,dtrain.knn$Safety.knn ,dtrain.knn$Value.knn)
dtest.knn2=data.frame(dtest.knn$Buying_price,dtest.knn$Maintenance_price.knn,dtest.knn$Lug_boot.knn,dtest.knn$Safety.knn ,dtest.knn$Value.knn)
dtrValue2=dtrain.knn2[,5]
dteValue2=dtest.knn2[,5]

results<-list()

for (k in 1:20) {
  set.seed(1)
knn.pred2=knn(dtrain.knn2[,1:4],dtest.knn2[,1:4],dtrValue2,k)
table.test.knn2=table (knn.pred2,dteValue2)
p2=sum(diag(table.test.knn2))/sum(table.test.knn2)
print(p2)
results=c(results,list(c(k,p2)))
  
}
q<-as.data.frame(results)
q<-t(q) #transpose q
q<-as.data.frame(q) #converts to data frame

#Renaming rows and columns
colnames(q)<-c("k","Precision")
rownames(q)<-c()

#Extracting the maximum precision and its associated K
qmax<-subset(q, (q$Precision==max(q$Precision)))
qmax

knn.pred2=knn(dtrain.knn2[,1:4],dtest.knn2[,1:4],dtrValue2,3)
table.test.knn2=table (knn.pred2,dteValue2)
table.test.knn2
p2=sum(diag(table.test.knn2))/sum(table.test.knn2)
p2
```
In this case, the highest precision is 69%. 


#<font color="blue">5) Post-analysis </font>

<h2><font color="teal">5.1) Summary </font></h2>

Among the 3 classifiers that we tested, KNN succeeds in reaching the highest precision (KNN:92.6%; LDA:86.7%; LR:65.7%) and we decide to choose KNN because of its accuracy of prediction.

<h2><font color="teal">5.2) Application of selected model on real examples </font></h2>

We used data from four real cars: the 911 sports car, the Audi A2, a Golf and a Twingo. We tested our KNN algorithm with 6 attributes and k=3.

```{r}
#Constitute the same training set as before

dtrain.knn.ex=data.frame(dtrain.knn$Buying_price,dtrain.knn$Maintenance_price.knn,dtrain.knn$Doors.knn,dtrain.knn$Persons.knn,dtrain.knn$Lug_boot.knn,dtrain.knn$Safety.knn,dtrain.knn$Value.knn)

#Import the excel containing the examples and forming the test data frame

examples<-read_xlsx("examples.xlsx")
examples<-as.data.frame(examples)
Safety.knn0=factor(examples$Safety, levels=c("low","med","high"),labels=c("0","1","2"))
Buying_price.knn0=factor(examples$Buying_price, levels=c("low","med","high","vhigh"),labels=c("0","1","2","4"))
Maintenance_price.knn0=factor(examples$Maintenance_price, levels=c("low","med","high","vhigh"),labels=c("0","1","2","4"))
Doors.knn0=factor(examples$Doors, levels=c("2","3","4","5more"),labels=c("0","1","2","3"))
Persons.knn0=factor(examples$Persons, levels=c("2","4","more"),labels=c("0","1","2"))
Lug_boot.knn0=factor(examples$Lug_boot, levels=c("small","med","big"),labels=c("0","1","2"))
Value.knn0=factor(examples$Value, levels=c("unacc","acc","good","vgood"),labels=c("1","2","3","4"))

dtest.knn.ex=data.frame(Buying_price.knn0,Maintenance_price.knn0,Doors.knn0,Persons.knn0,Lug_boot.knn0,Safety.knn0,Value.knn0)


dtrValue=dtrain.knn.ex[,7]
dteValue=dtest.knn.ex[,7]

#Apply our model to examples set

set.seed(1)
knn.pred.ex=knn(dtrain.knn.ex,dtest.knn.ex,dtrValue,k=3)
table.test.knn.ex=table (knn.pred.ex,dteValue)
table.test.knn.ex
p0=sum(diag(table.test.knn.ex))/sum(table.test.knn.ex)
#Precision on our examples set 
p0

heatmap=heatmap.2(table.test.knn0,col=rev(heat.colors(100)),xlab="Test_set",ylab="Prediction",cellnote=table.test.knn.ex,notecol="black",trace="none",dendrogram="none",rownames(table.test.knn.ex),colnames(table.test.knn.ex))
CrossTable(dteValue,knn.pred.ex,prop.chisq = FALSE)
```
The precision on this examples set is 75%.
